---
title: "analyisis_la_r01"
author: "Nick R. Bachelder"
date: '2022-07-13'
output: html_document
---

/Users/nickbachelder/Desktop/Duke Internship 2021/LA_Cohort_R01/LA_A1C

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align="center")

setwd(here())

load(file = "analysis_prep.RData")

library(dplyr)
library(ggplot2)
library(caret)
library(tidyr)
library(hrbrthemes)
library(stringr)
library(randomForest)
require(dplyr)
library(MASS)
library(knitr)
library(kableExtra)
library(mvtnorm)
library(scales)
library(DT)
library(readxl)
library(reshape2)
library(zoo)
library(ggcorrplot)
library(psych)
library(tidyverse)
library(gtsummary)
library(gtExtras)
library(bstfun)
library(MplusAutomation)
library(tidyLPA)
library(gridExtra)
```


## Data Cleaning Comments:
* Selected variables include **all cytokines, both CHiP variables, ZIP codes, the food desert variable**, and those **survey variables** which were convenient for modeling. All variables are summarized below.
* Dietary weekly consumption variables (ie. servings of oil, tacos, etc) were grouped into more broad catagories (Surgar, Red Meats, Fatty Carbs, Dairy, Breads)

  * **fatty carbs**: pizza, fries, chips, takeout, pastries, cake, pancakes, kfc, tacos
  * **red meats**: mince, beef, sausages, bacon
  * **high sugar products**: sports_drinks, soda, sugar, ice_cream, chocolate, lollies, spreads
  * **vitamin**: vitamin
  * **bread**: bread, rice, tortilla
  * **dairy**: eggs, cheese, milk
  * **remove** due to broad health range: -dressings, -beans, -oil

* Food deserts were recorded at census level and patient data was at the zip code level. Since zip codes are larger than census tracts, food desert variables had to to be summarized to be the **percent of zip code population that lives in food desert conditions**.
* Survey variables that recorded how patients transported to get food (ie. Car, Bus, Metro, Walk, Bike) were summarized more broadly into three catagories (Car N = 296, Car & Other N = 33, and No Car N = 5)
* While data contained ZIP+4 values, there is no more detailed translation from the zip+4 to census tract. Therefore, ZIP was simplified to the more simple 5-digit ZIP
* The same definition for food desert was used as the food desert poster (>1/2 for urban and >10 for rural)
* Box-cox transformations was used on ALL variables are shapiro-wilk tests accepted non-normality alternative hypothesis for all variables.
* Only select models were tests due to (1) scatter plots with A1C failing to show linear relationships needed to satisfy assumptions of linear models and (2) overlap of patients between CHiP data, survey data, and ZIP code included data was limited, with fully complete patient observations being only n= 11 out of total N = 411.


## Summarize Variables:
* Looks as though many variables are skewed and non-normal.
```{r}
cohort_df %>% dplyr::select(-c(men_status, ethnicity)) %>%
  tbl_summary(missing = "ifany",
              missing_text = "N missing",
              type = list(all_continuous() ~ "continuous2", housemates ~ "continuous2", 
                          income_status ~ "continuous2", education_status ~ "continuous2", health_status ~ "continuous2"),
              statistic = all_continuous() ~ c("{median} ({p25}, {p75})",
                                               "({min}, {max})")) %>%
  
  bstfun::add_sparkline(type = "histogram")
```

## Dietary LCA:
* To see the profiles of dietary groups, LCA was attempted on the 5 catagories of food consumption.
* Groups were best divided into 3 latent groups (by lowest BIC), and it seems that each class eats progressively more servings of everything. This raises questions about whether or not survey diet data is accurate, or if different individuals have different ideas of serving scales, causing such variation and latent group consistancy.
* To check whether survey data is actually accurate, multiple hypothesis testing was preformed on each latent group with the idea that the groups that consumed more should be significantly associated with higher weight and groups that consumed less with lower weight. Such **hypothesis tests showed latent class association with A1C, weight, and BMI**.

```{r}
df_lca <- cohort_df %>%
  filter(complete.cases(weekly_dairy, bread_weekly, high_sugar_weekly, red_meats_weekly, fatty_carbs_weekly))


results_lca <- df_lca %>% dplyr::select(weekly_dairy, bread_weekly, high_sugar_weekly, red_meats_weekly, fatty_carbs_weekly) %>%
    estimate_profiles(1:5,
                      package = 'MplusAutomation',
                      variances = c("varying"),
                      covariances = c('varying'))

get_data(results_lca) %>% filter(classes_number == 3, Class == Class_prob) %>% group_by(Class) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  dplyr::select(Class, weekly_dairy, bread_weekly, high_sugar_weekly, red_meats_weekly, fatty_carbs_weekly) %>% 
  pivot_longer(cols = c(weekly_dairy, bread_weekly, high_sugar_weekly, red_meats_weekly, fatty_carbs_weekly)) %>% 
  ggplot(aes(x = name, y = value, group = Class, fill = as.factor(Class))) + geom_bar(position = 'dodge', stat = 'identity') + 
  ggtitle('Estimated Mean Parameters for Each Dietary Latent Cluster') + scale_fill_discrete(name = 'Latent Cluster') + 
  ylab('Mean Serving Consumption Weekly') + xlab('Spatial Outcome') + theme(text = element_text(size=9)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```




```{r}
lca_corr <- get_data(results_lca) %>% filter(classes_number == 3, Class == Class_prob) %>% 
  dplyr::select(Class, weekly_dairy, bread_weekly, high_sugar_weekly, red_meats_weekly, fatty_carbs_weekly)

lca_corr <- df_lca %>% left_join(lca_corr, by = c('weekly_dairy', 'bread_weekly', 'high_sugar_weekly', 'red_meats_weekly', 'fatty_carbs_weekly')) %>% 
  dplyr::select(HgA1c, weight, bmi, Class) %>% filter(complete.cases(.)) %>%
  mutate(Class = as.factor(Class))


tibble(outcome = c("weight", "HgA1c", "bmi")) %>%
  rowwise() %>%
  mutate(
    tbl = 
      lm(str_glue("{outcome} ~ Class"), data = lca_corr) %>%
      tbl_regression() %>%
      add_significance_stars(
        hide_se = TRUE,
        hide_ci = FALSE
      ) %>%
      list()
  ) %>%
  # pull tbl_regression() objects into single merged table
  pull(tbl) %>%
  tbl_merge(tab_spanner = c("**Weight**", "**A1C**", "**BMI**")) %>%
  # add table captions
  as_gt() %>%
  gt::tab_header(title = "Table 1. Latent Class Association")
```
## ACS LCA:
* To see the profiles of dietary groups, LCA was attempted on the advantage and disadvantage variables as done in previous work.
* Association with advantage latent class is almost significant. 

```{r}
get_data(results_lca_DIS) %>% filter(classes_number == 2, Class == Class_prob) %>% group_by(Class) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  dplyr::select(Class, crowded_prop.T, female_householder_prop.T, pop_non_hispanic_black_prop.T, vehicle_none_prop.T, no_internet_prop.T, public_assistance_prop.T,
                housing_rental_prop.T,poverty_prop.T, single_prop.T) %>% 
  pivot_longer(cols = c(crowded_prop.T, female_householder_prop.T, pop_non_hispanic_black_prop.T, vehicle_none_prop.T, no_internet_prop.T, public_assistance_prop.T,
                housing_rental_prop.T,poverty_prop.T, single_prop.T)) %>% 
  ggplot(aes(x = name, y = value, group = Class, fill = as.factor(Class))) + geom_bar(position = 'dodge', stat = 'identity') + 
  ggtitle('Estimated Mean Parameters for Each Disadvantage Latent Cluster') + scale_fill_discrete(name = 'Latent Cluster') + 
  ylab('Mean') + xlab('Spatial Outcome') + theme(text = element_text(size=9)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

get_data(results_lca_ADV) %>% filter(classes_number == 2, Class == Class_prob) %>% group_by(Class) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  dplyr::select(Class, fem_mgmt_prof_prop.T, male_mgmt_prof_prop.T, hs_grad.T) %>% 
  pivot_longer(cols = c(fem_mgmt_prof_prop.T, male_mgmt_prof_prop.T, hs_grad.T)) %>% 
  ggplot(aes(x = name, y = value, group = Class, fill = as.factor(Class))) + geom_bar(position = 'dodge', stat = 'identity') + 
  ggtitle('Estimated Mean Parameters for Each Advantage Latent Cluster') + scale_fill_discrete(name = 'Latent Cluster') + 
  ylab('Normalized Mean') + xlab('Spatial Outcome') + theme(text = element_text(size=9)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

modeling_df <- compare_df_tranformed
  
modeling_df <- modeling_df %>% dplyr::select(HgA1c.T , Class_ADV, Class_DIS)

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

mod4 <- lm(HgA1c.T ~ ., modeling_df)
mod4 %>% tbl_regression(estimate_fun = purrr::partial(style_ratio, digits = 4))
```

## Cytokine LCA

```{r}
get_data(results_cyto_lca) %>% filter(classes_number == 4, Class == Class_prob) %>% group_by(Class) %>%
  dplyr::summarise(across(everything(), mean)) %>% 
  dplyr::select(Class, TNF_alpha.T, IL_6.T, Leptin.T, IL_15.T, IL_23.T, CXCL1.T) %>% 
  mutate(across(c('TNF_alpha.T', 'IL_6.T', 'Leptin.T', 'IL_15.T', 'IL_23.T', 'CXCL1.T'),   function(x) {scale(x, center = TRUE, scale = TRUE) })) %>%
  pivot_longer(cols = c(TNF_alpha.T, IL_6.T, Leptin.T, IL_15.T, IL_23.T, CXCL1.T)) %>% 
  ggplot(aes(x = name, y = value, group = Class, fill = as.factor(Class))) + geom_bar(position = 'dodge', stat = 'identity') + 
  ggtitle('Estimated Mean Parameters for Each Cytokine Latent Cluster') + scale_fill_discrete(name = 'Latent Cluster') + 
  ylab('Normalized Mean') + xlab('Spatial Outcome') + theme(text = element_text(size=9)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

modeling_df <- compare_df_tranformed
  
modeling_df <- modeling_df %>% dplyr::select(HgA1c.T , Class_CYTO)

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

mod4 <- lm(HgA1c.T ~ ., modeling_df)
mod4 %>% tbl_regression(estimate_fun = purrr::partial(style_ratio, digits = 4))
```



## Heatmap correlations of survey data and heatmap associations of cytokines, CHiP, and food desert.
```{r}
ggcorrplot(cor_df, hc.order = TRUE, type = "lower",
     outline.col = "white")
ggcorrplot(cor_surv, hc.order = TRUE, type = "lower",
     outline.col = "white")
```

# Check significance of above correlations. Yellow is significant and red is extremely significant.
```{r}
combined %>% 
  mutate_all(~cell_spec(.x, background = case_when(grepl('⋆⋆⋆⋆', .x) ~ "red",
                                                   grepl('⋆⋆⋆', .x) ~ "red",
                                                   grepl('⋆⋆', .x) ~ "yellow",
                                                   grepl('⋆', .x) ~ "yellow",
                                                   TRUE ~ "white"))) %>%
  kbl(booktabs = T, escape = F, caption = "Correlation, Significance, and Sample Size For HgA1c, Cytokine, Survey, and Food Desert Values") %>% kable_classic_2() %>%
      kable_paper(full_width = F) %>%
  footnote(symbol = c("Pearson Correlation is used", "Holm-Bonferroni method is used to compute p-values for multiple comparisons"),
           footnote_as_chunk = T) 
```

## For modeling, check the distributions of dependent and A1C variables.
* Seem non-normal and non-symetric, consider transformation.

```{r}
combine_df_non_ACS <- combine_df1[,1:34]
combine_df_ACS <- combine_df1[,35:ncol(combine_df1)]

l <- list()
i = 1
for (value in names(combine_df_non_ACS %>% select_if(is.numeric))) {
  l[[i]] <-  ggplot(data = combine_df_non_ACS, aes_string(x = value)) + geom_histogram()
  i <- i+1
}

n <- length(l)
nCol <- floor(sqrt(n))
do.call("grid.arrange", c(l, ncol=nCol))
```

```{r}
l <- list()
i = 1
for (value in names(combine_df_ACS %>% select_if(is.numeric))) {
  l[[i]] <-  ggplot(data = combine_df_ACS, aes_string(x = value)) + geom_histogram()
  i <- i+1
}

n <- length(l)
nCol <- floor(sqrt(n))
do.call("grid.arrange", c(l, ncol=nCol))
```


## Check normality of dependent variables using Wilk-Shapiro tests.
***  All variables are non-normal. We will box transform.
```{r}
## check for normality of each variable:
get_value_shapiro <- function(x) {
  shapiro.test(x)$p}
data.frame(unlist(lapply(numeric_final, get_value_shapiro))) %>% `colnames<-`('p_value') %>% 
  mutate(p_value = case_when(p_value < 0.05 ~ paste('Non-Normal (p:', as.character(round(p_value, 7)), ')'),
                             TRUE ~ paste('Normal: (p:', as.character(round(p_value, 7)), ')'))
         ) %>% kbl(booktabs = T, escape = F, caption = "Wilks-Shapiro Normality") %>% kable_classic_2() %>%
      kable_paper(full_width = F)
```


## Natural log tranform for non-ACS, square root for ACS - now recheck for data symmetry
* Looks much more symetric and closer to normal

```{r}
combine_df_non_ACS_T <- numeric_transformed[,1:23]
combine_df_ACS_T <- numeric_transformed[,34:ncol(numeric_transformed)]

l <- list()
i = 1
for (value in names(combine_df_non_ACS_T)) {
  l[[i]] <-  ggplot(data = combine_df_non_ACS_T, aes_string(x = value)) + geom_histogram()
  i <- i+1
}

n <- length(l)
nCol <- floor(sqrt(n))
do.call("grid.arrange", c(l, ncol=nCol))
```

```{r}
l <- list()
i = 1
for (value in names(combine_df_ACS_T)) {
  l[[i]] <-  ggplot(data = combine_df_ACS_T, aes_string(x = value)) + geom_histogram()
  i <- i+1
}

n <- length(l)
nCol <- floor(sqrt(n))
do.call("grid.arrange", c(l, ncol=nCol))
```

## Check for linear association among variables to select modeling variables
* Potential continuous variables with linear relationship with AIC inlude **CXCL1, Food Desert Percentage, Health Status, IL6_K9ac_ChIP, hs_grad and TNFa_K9ac_Chip**. We will also attempt to include factor variables **family_w_diabetes, Race, and Transportation_type_for_food, and ACS latent classes**
```{r}
combine_df_non_ACS_T <- numeric_transformed[,1:23]
combine_df_ACS_T <- numeric_transformed[,c(13, 34:ncol(numeric_transformed))]

l <- list()
i = 1
for (value in names(combine_df_non_ACS_T)) {
  l[[i]] <-  combine_df_non_ACS_T %>% ggplot(aes_string(x = value, y = 'HgA1c.T')) + geom_point()
  i <- i+1
}

n <- length(l)
nCol <- 5
do.call("grid.arrange", c(l, ncol=nCol))
```

```{r}
l <- list()
i = 1
for (value in names(combine_df_ACS_T)) {
  l[[i]] <-  combine_df_ACS_T %>% ggplot(aes_string(x = value, y = 'HgA1c.T')) + geom_point()
  i <- i+1
}

n <- length(l)
nCol <- 5
do.call("grid.arrange", c(l, ncol=nCol))
```

## Issue: 
* There are only 11 patients with interesting selected continuous variables (CXCL1, Food Desert Percentage, Health Status, IL6_K9ac_ChIP, and TNFa_K9ac_Chip) with no missing values. We will need to break up models. 

  * Model 1: A1C ~ Percent_of_tract_in_FD.T + CXCL1 + family_w_diabetes + Transportation_type_for_food (N = 94)
  * Model 2: A1C ~ IL6_K9ac_ChIP + TNFa_K9ac_Chip (N = 17)
  * Model 3: A1C ~ us_born + Race (N = 155)
  * Model 4: A1C ~ All ACS variables (N = 165)
  * Model 5: A1C ~ TNF_alpha.T + Leptin.T (N = 154)
  * **NOTE: For better modeling, maybe we can discuss getting updated cytokynes measures for patients to increase sample sizes**

## Model 1: 
* Percent of Tract in Food Desert and Family w Diabetes are significant despite small coefficients. 

```{r}
library(purrr)
modeling_df <-compare_df_tranformed
  
modeling_df <- modeling_df %>% dplyr::select(HgA1c.T, Percent_of_tract_in_FD.T, CXCL1.T, family_w_diabetes, 
                                             Transportation_type_for_food)

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

mod4 <- lm(HgA1c.T ~ ., modeling_df)
mod4 %>% tbl_regression(estimate_fun = purrr::partial(style_ratio, digits = 4))
```
## Model 2: 
* TNFa_K9ac_ChIP significant
* As a note, there is strong positive correlation between the two chip variables, so consider elaster net on this model to reduce multicollinearity concern and introduce regularization.

```{r}
modeling_df <- numeric_transformed %>% dplyr::select(HgA1c.T, IL6_K9ac_ChIP.T, TNFa_K9ac_ChIP.T)

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

mod4 <- lm(HgA1c.T ~ ., modeling_df)
mod4 %>% tbl_regression(estimate_fun = purrr::partial(style_ratio, digits = 3))
```


## Model 3: 
* None significant in this cohort

```{r}
modeling_df <- compare_df_tranformed %>% dplyr::select(HgA1c.T, us_born, Race)

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

mod4 <- lm(HgA1c.T ~ Race + us_born, modeling_df)
mod4 %>% tbl_regression(estimate_fun = purrr::partial(style_ratio, digits = 3))
```

## Model 4: 
* None independently significant despite significance of latent class for ACS

```{r}
modeling_df <- combine_df_ACS_T

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

mod4 <- lm(HgA1c.T ~ ., modeling_df)
mod4 %>% tbl_regression(estimate_fun = purrr::partial(style_ratio, digits = 3))
```


# Model 5
*None-Significant as implied by scatter plots and lack of linear trends

```{r}
modeling_df <- numeric_transformed
  
modeling_df <- modeling_df %>% dplyr::select(HgA1c.T , bmi.T , weight.T , health_status.T , TNF_alpha.T , IL_6.T , Leptin.T)

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

mod4 <- lm(HgA1c.T ~ Leptin.T + TNF_alpha.T, modeling_df)
mod4 %>% tbl_regression(estimate_fun = purrr::partial(style_ratio, digits = 4))
```




## Elastic Net Model 2: 
* Both CHiPs are selected in elastic net model, but both coeficients are shrunk to be very small compared to the intercept.

```{r}
library(glmnet)

modeling_df <- numeric_transformed %>% dplyr::select(HgA1c.T, IL6_K9ac_ChIP.T, TNFa_K9ac_ChIP.T)

modeling_df <- modeling_df[complete.cases(modeling_df),] %>% data.frame()

## Create modeling data
X <- modeling_df 

X2 <-  model.matrix(HgA1c.T ~ ., X)
Y2 <- X$HgA1c.T


# Create cv results df
df1 = data.frame(
    enum = integer(),
    alpha = double(),
    cvm.min = double(),
    cvm.1se = double(),
    lambda.min = double(),
    lambda.1se = double())

  # - select alpha and lambda via CV
enum = 0
for (alpha in c(0, 0.01, 0.1, 0.5, 1)) {
  enum <-  enum + 1
  fit <-  cv.glmnet(X2, Y2, family='gaussian', alpha=alpha, nfolds = 3)
  cvm.min <-  fit$cvm[fit$lambda == fit$lambda.min]
  cvm.1se <-  fit$cvm[fit$lambda == fit$lambda.1se]
  lambda.min <-  fit$lambda.min
  lambda.1se <-  fit$lambda.1se
  
  assign(paste("fit", enum, sep=""), fit)
  
  df1_temp <-  data.frame(enum, alpha, cvm.min, cvm.1se, lambda.min, lambda.1se)
  
  df1 <-  rbind(df1, df1_temp) }

# - select model that minimizes CV error
best.model <-  df1[df1$cvm.min==min(df1$cvm.min),]
best.fit   <-  get(paste('fit', best.model$enum, sep=''))

# extract non-zero coefficients from best model
coef  <-  coef(best.fit, s=best.fit$lambda.min) ## uses lowest mse lambda
coef2 <- coef(best.fit, s=best.fit$lambda.1se) ## uses 1 se larger from lowest mse lambda (recomended for more regularization)

coef <-  data.frame(
  vars   = coef@Dimnames[[1]][ which(coef != 0 ) ], 
  val     = coef              [ which(coef != 0 ) ] )


plot_min <- coef %>%
  dplyr::rename('row' = 'vars', 'value' = 'val') %>%
  ggplot(aes(value, reorder(row, value), color = value > 0.001)) +
  geom_point(show.legend = FALSE) +
  ggtitle("Influential variables") +
  xlab("Coefficient") +
  ylab(NULL)

most_sig <- coef %>% filter(abs(val) %in% tail(sort(abs(coef$val)), 10) )

plot_most_sig <- most_sig %>%
  dplyr::rename('row' = 'vars', 'value' = 'val') %>%
  ggplot(aes(value, reorder(row, value), color = value > 0)) +
  geom_point(show.legend = FALSE) +
  ggtitle("Non-Zero Coeficients in Elastic Net Model") +
  xlab("Coefficient") +
  ylab(NULL)

coef <- coef  %>% mutate(val = round(val, 6)) 

return_this <- c()

return_this$df1 <- df1
return_this$plot_min <- plot_min
return_this$coef <- coef %>% arrange(desc(val))

return_this$plot_min
```


## Food desert vs A1C map 
* Across LA county

```{r, include =  FALSE}
library(rgdal)
library(rgeos)
shape <- readOGR('/Users/nickbachelder/Desktop/Duke Internship 2021/LA_Cohort_R01/tl_2021_us_zcta520', layer = 'tl_2021_us_zcta520')

plot_df <- combine_df1 %>% dplyr::select(Curr_ZIP, Percent_of_tract_in_FD, HgA1c) %>% group_by(Curr_ZIP) %>% 
  dplyr::summarize(Percent_of_tract_in_FD = max(Percent_of_tract_in_FD), HgA1c_mean = mean(HgA1c))

zips_use <- read_csv('/Users/nickbachelder/Desktop/Duke Internship 2021/LA_Cohort_R01/Zip_Codes_(LA_County).csv')$ZIP # LA county


shape@data$id <- rownames(shape@data)


shape <- shape[shape$ZCTA5CE20 %in% zips_use,]



shape@data <- left_join(shape@data %>% dplyr::rename('Curr_ZIP' = 'ZCTA5CE20'), plot_df, by = 'Curr_ZIP')
```


```{r}
library(viridis)
shp_df <- fortify(shape)
shp_df   <- left_join(shp_df %>% mutate(id = as.numeric(id)), shape@data %>% mutate(id = as.numeric(id)), by="id")

plot_food_desert <- ggplot() + geom_polygon(data = shp_df, aes(x = long, y = lat, group = group, fill = Percent_of_tract_in_FD)) + theme_void() +
  theme(legend.position="none") +
  ggtitle('Food Desert Percentage')


plot_A1C <- ggplot() + geom_polygon(data = shp_df, aes(x = long, y = lat, group = group, fill = HgA1c_mean)) + theme_void() +
  theme(legend.position="none") + ggtitle('Mean A1C Level')

grid.arrange(plot_food_desert, plot_A1C, ncol=2)
```




